One-armed spirals in locally isothermal, radially structured self-gravitating discs

### 2014-11-16 paper/results_3d.tex
updating 3d stuff

The 3D disc has radial size
$[r_\mathrm{min},r_\mathrm{max}]=[0.4,10]R_0$ and vertical extent  
$n_H=2$ scale-heights at $R=R_0$. The resolution is $N_r\times N_\theta\times
N_\phi=256\times32\times256$. This corresponds to about $4$ cells per
$H$ in radius. Because of the much reduced resolution 
compared to 2D, we use a smooth perturbation by setting
$\delta = 10^{-3}$ and $M=1$ in Eq. \ref{randpert}. This corresponds
to a single $m=1$ disturbance in $R\in[R_1,R_2]$, which was found to dominate
the 2D simulation.

The 3D discs are initialised in approximate equilibrium only, so we
first evolve the disc without perturbations using  
$(\lmax,\mmax)=(32,0)$ up to $t=10P_0$, during which 
meridional velocities are damped out. We then restart the simulation
with the above perturbation and $(\lmax,\mmax)=(32,32)$ and damp
merdional velocities near the radial boundaries.

Fig. \ref{3d_ampmax} plots the amplitude of the $m=1$ spirals measured
in the ZEUS-MP and PLUTO runs. 
%why no high m
                                %modes as in 2D for iso disc?
                                %resolution? 
In ZEUS-MP we observe disturbances developed near
the inner boundary, which is probably responsible for the initial growth
for $t<70P_0$. This is likely a numerical artifact and effectively seeds
the ZEUS-UP simulation with a larger initial perturbation. Results
from ZEUS-MP are therefore off-set from PLUTO. However,
once the coherent $m=1$ spiral begins to grow ($t\gtrsim 100P_0$), 
we measure similar growth rates in both codes: 
\begin{align*}
  &\gamma \simeq 0.0073\Omega_k(R_0) \quad\quad \mathrm{PLUTO},\\
  &\gamma \simeq 0.0084\Omega_k(R_0) \quad\quad \varA{ZEUS-MP}.
\end{align*}
Both are smaller than the 2D simulations and is probably due to the
low resolutions adopted and/or the small softening length employed in
2D.  Fig. \ref{3d_ampmax} also show results from strictly isothermal
simulations, which display no growth compared to cases with a 
temperature gradient. This confirms the temperature gradient effect is
the same in 3D.

@@

\begin{figure}
  \includegraphics[width=\linewidth]{figures/m1_analysis_plot_ampmax3d.ps}
  \caption{Evolution of the maximum $m=1$ density in  $r\in[R_1,R_2]$
    measured in the 3D simulations. Results from discs with a imposed 
    temperature gradient ($q=1$) and a strictly isothermal disc
    ($q=0$) are shown.  
    \label{3d_ampmax}}   
\end{figure}

>>>>

The 3D disc has radial size
$[r_\mathrm{min},r_\mathrm{max}]=[0.4,10]R_0$ and vertical extent  
$n_H=2$ scale-heights at $R=R_0$. The resolution is $N_r\times N_\theta\times
N_\phi=256\times32\times512$. This corresponds to about $4$ cells per
$H$ in radius. Because of the reduced resolution 
compared to 2D, we use a smooth perturbation by setting
$\delta = 10^{-3}$ and $M=1$ in Eq. \ref{randpert}. This corresponds
to a single $m=1$ disturbance in $R\in[R_1,R_2]$, which
eventually dominates in 2D simulations.

The 3D discs are initialised in approximate equilibrium only, so we
first evolve the disc without perturbations using  
$(\lmax,\mmax)=(32,0)$ up to $t=10P_0$, during which 
meridional velocities are damped out. We then restart the simulation
with the above perturbation and $(\lmax,\mmax)=(32,32)$, and damp
merdional velocities near the radial boundaries.

Fig. \ref{3d_ampmax} plots the amplitude of the $m=1$ spirals measured
in the ZEUS-MP and PLUTO runs.  
%why no high m
                                %modes as in 2D for iso disc?
                                %resolution? 
In the ZEUS-MP run we observe disturbances developed near
the inner boundary, which is probably responsible for the initial growth
for $t<70P_0$. This is likely a numerical artifact and effectively seeds
the ZEUS-UP simulation with a larger initial perturbation. Results
from ZEUS-MP are therefore off-set from PLUTO by $\sim50P_0$ However,
once the coherent $m=1$ spiral begins to grow ($t\gtrsim 100P_0$), 
we measure similar growth rates in both codes: 
\begin{align*}
  &\gamma \simeq 0.0073\Omega_k(R_0) \quad\quad \mathrm{PLUTO},\\
  &\gamma \simeq 0.0084\Omega_k(R_0) \quad\quad \varA{ZEUS-MP}.
\end{align*}
Both are smaller than the 2D simulations and due to the
low resolutions adopted and/or the small softening length employed in
2D.  Fig. \ref{3d_ampmax} also show results from strictly isothermal
simulations, which display no growth compared to cases with a 
temperature gradient. This confirms the temperature gradient effect is
the same in 3D.

@@

\begin{figure}
  \includegraphics[width=\linewidth]{figures/m1_analysis_plot_ampmax3d.ps}
  \caption{Evolution of the maximum $m=1$ density component in  $r\in[R_1,R_2]$
    measured in the 3D simulations. Results from discs with a imposed 
    temperature gradient ($q=1$) and a strictly isothermal disc
    ($q=0$) are shown.  
    \label{3d_ampmax}}   
\end{figure}
